Hydrodynamics and rheology of a vesicle doublet suspension

### 2019-07-09 vesicle_code/docs/adhesion/adhesion.tex
worked on Figure 12 (left)
added extensional plots

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Adhesion of two vesicles in an extensional flow} 
\label{sec:eflow} 
The hydrodynamics of a single vesicle in an extensional flow has
revealed novel nonlinear vesicle dynamics not found for a viscous
drop~\cite{KantslerSegreSteinberg2008_PRL, ZhaoShaqfeh2013_JFM,
Narsimhan2014_JFM, DahlNarsimhanGouveia2016_SoftMatt}.  The stagnation
point (zero-velocity at the origin) in an extensional flow, $\uu
=\chi(-x,y)$, is a saddle point where two laminar streams converge along
an axis and then diverge in the orthogonal direction.  With an active
control algorithm to place the stagnation point at a desirable location
by adjusting the streaming flow strength with a feedback
loop~\cite{Johnson-Chavarria2011_EMJ}, a particle can be trapped at the
stagnation point for long time scales to facilitate image acquisition or
other detailed measures such as particle image velocimetry of flow
inside and around the particle.  To explore the application of such a
fluid trap to measure the adhesion strength between two bound vesicle
membranes, we propose the following experiment.  Beginning with two
identical vesicles in an equilibrium configuration that form a doublet
with a flat contact region (Figure~\ref{fig:Dec18_vesicle_shape}), we
turn on the fluid trap with the stagnation point placed at the center of
the vesicle doublet.  Depending on the Hamaker constant, we expect that
either the flow overcomes the adhesive force and the doublet is broken,
or the adhesive force is sufficiently strong and the doublet reaches a
stable stationary configuration.

@@

In Figure~\ref{fig:sflow_distance}(b), we plot the time required for the
doublet to make a complete revolution at a variety of Hamaker constants
and shear rates.  For a fixed Hamaker constant, as expected, the period
decreases for larger shear rates.  In addition, for a fixed shear rate,
the period decreases as the Hamaker constant increases, but the period
eventually plateaus.  The dynamics of these doublet have similar
behavior to a single rigid body undergoing a Jeffery orbit in a shear
flow.  Therefore, we define the effective aspect ratio
$\lambda_{\mathrm{eff}}$ which satisfies
\begin{align*}
  \frac{2\pi}{\chi}\left(\lambda + \lambda^{-1} \right) = p,
\end{align*}
where $p$ is the period of the doublet.  The effective aspect ratio is
the aspect ratio of an ellipse in a shear flow that is undergoing a
Jeffery orbit with period $p$.  For the Hamaker constant $\mathcal{H} =
2$, the shear rates, periods, and effective aspect ratio are summarized
in Table~\ref{tbl:aspectRatio}.  Since the effective aspect ratio is
comparable for all three shear rates, this indicates that the dynamics
of a vesicle doublet in a shear flow resembles a rigid body undergoing a
Jeffery orbit.
\begin{table}[htp]
  \begin{tabular}{>{\centering\arraybackslash}p{2cm}
>{\centering\arraybackslash}p{2cm}
>{\centering\arraybackslash}p{2cm}}
  $\chi$ & $p$ & $\lambda_{\mathrm{eff}}$ \\
  \hline
  0.50 & 38.06 & 2.92 \\
  0.75 & 24.81 & 2.85 \\
  1.00 & 19.47 & 3.00 
\end{tabular}
\caption{\label{tbl:aspectRatio} The shear rate, period, and effective
aspect ratio of a vesicle doublet with reduced area $\Delta A = 0.9$,
Hamaker constant $\mathcal{H} = 2$, and separation distance $\delta =
0.4$.  The effective aspect ratios are comparable for all three shear
rates.}
\end{table}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Adhesion of two vesicles in an extensional flow} 
\label{sec:eflow}

\begin{figure}[htp]
  \includegraphics[width = 0.16\textwidth,trim={4cm 0.5cm 4cm 0.5cm},clip]
    {figs/extensional_displacement0_adS7em1_image01.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 0.5cm 4cm 0.5cm},clip]
    {figs/extensional_displacement0_adS7em1_image02.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 0.5cm 4cm 0.5cm},clip]
    {figs/extensional_displacement0_adS7em1_image03.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 0.5cm 4cm 0.5cm},clip]
    {figs/extensional_displacement0_adS7em1_image04.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 0.5cm 4cm 0.5cm},clip]
    {figs/extensional_displacement0_adS7em1_image05.png} \\

  \includegraphics[width = 0.16\textwidth,trim={4cm 0.5cm 4cm 0.5cm},clip]
    {figs/extensional_displacement2em2_adS7em1_image01.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 0.5cm 4cm 0.5cm},clip]
    {figs/extensional_displacement2em2_adS7em1_image02.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 0.5cm 4cm 0.5cm},clip]
    {figs/extensional_displacement2em2_adS7em1_image03.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 0.5cm 4cm 0.5cm},clip]
    {figs/extensional_displacement2em2_adS7em1_image04.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 0.5cm 4cm 0.5cm},clip]
    {figs/extensional_displacement2em2_adS7em1_image05.png} \\

  \includegraphics[width = 0.16\textwidth,trim={4cm 0.5cm 4cm 0.5cm},clip]
    {figs/extensional_displacement1em1_adS7em1_image01.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 0.5cm 4cm 0.5cm},clip]
    {figs/extensional_displacement1em1_adS7em1_image02.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 0.5cm 4cm 0.5cm},clip]
    {figs/extensional_displacement1em1_adS7em1_image03.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 0.5cm 4cm 0.5cm},clip]
    {figs/extensional_displacement1em1_adS7em1_image04.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 0.5cm 4cm 0.5cm},clip]
    {figs/extensional_displacement1em1_adS7em1_image05.png} \\

  \includegraphics[width = 0.16\textwidth,trim={4cm 0.5cm 4cm 0.5cm},clip]
    {figs/extensional_displacement1em1_adS6em1_image01.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 0.5cm 4cm 0.5cm},clip]
    {figs/extensional_displacement1em1_adS6em1_image02.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 0.5cm 4cm 0.5cm},clip]
    {figs/extensional_displacement1em1_adS6em1_image03.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 0.5cm 4cm 0.5cm},clip]
    {figs/extensional_displacement1em1_adS6em1_image04.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 0.5cm 4cm 0.5cm},clip]
    {figs/extensional_displacement1em1_adS6em1_image05.png}

  \caption{\label{fig:extensionalNonsymmetric} }
  \end{figure}

The hydrodynamics of a single vesicle in an extensional flow has
revealed novel nonlinear vesicle dynamics not found for a viscous
drop~\cite{KantslerSegreSteinberg2008_PRL, ZhaoShaqfeh2013_JFM,
Narsimhan2014_JFM, DahlNarsimhanGouveia2016_SoftMatt}.  The stagnation
point (zero-velocity at the origin) in an extensional flow, $\uu
=\chi(-x,y)$, is a saddle point where two laminar streams converge along
an axis and then diverge in the orthogonal direction.  With an active
control algorithm to place the stagnation point at a desirable location
by adjusting the streaming flow strength with a feedback
loop~\cite{Johnson-Chavarria2011_EMJ}, a particle can be trapped at the
stagnation point for long time scales to facilitate image acquisition or
other detailed measures such as particle image velocimetry of flow
inside and around the particle.  To explore the application of such a
fluid trap to measure the adhesion strength between two bound vesicle
membranes, we propose the following experiment.  Beginning with two
identical vesicles in an equilibrium configuration that form a doublet
with a flat contact region (Figure~\ref{fig:Dec18_vesicle_shape}), we
turn on the fluid trap with the stagnation point placed at the center of
the vesicle doublet.  Depending on the Hamaker constant, we expect that
either the flow overcomes the adhesive force and the doublet is broken,
or the adhesive force is sufficiently strong and the doublet reaches a
stable stationary configuration.

@@

In Figure~\ref{fig:sflow_distance}(b), we plot the time required for the
doublet to make a complete revolution at a variety of Hamaker constants
and shear rates.  For a fixed Hamaker constant, as expected, the period
decreases for larger shear rates.  In addition, for a fixed shear rate,
the period decreases as the Hamaker constant increases, but the period
eventually plateaus.  The dynamics of these doublet have similar
behavior to a single rigid body undergoing a Jeffery orbit in a shear
flow.  Therefore, we define the effective aspect ratio
$\lambda_{\mathrm{eff}}$ which satisfies
\begin{align*}
  \frac{2\pi}{\chi}\left(\lambda + \lambda^{-1} \right) = p,
\end{align*}
where $p$ is the period of the doublet.  The effective aspect ratio is
the aspect ratio of an ellipse in a shear flow that is undergoing a
Jeffery orbit with period $p$.  For the Hamaker constant $\mathcal{H} =
2$, the shear rates, periods, and effective aspect ratio are summarized
in Table~\ref{tbl:aspectRatio}.  Since the effective aspect ratio is
comparable for all three shear rates, this indicates that the dynamics
of a vesicle doublet in a shear flow resembles a rigid body undergoing a
Jeffery orbit.
\begin{table}[htp]
  \begin{tabular}{>{\centering\arraybackslash}p{2cm}
>{\centering\arraybackslash}p{2cm}
>{\centering\arraybackslash}p{2cm}}
  $\chi$ & $p$ & $\lambda_{\mathrm{eff}}$ \\
  \hline
  0.50 & 38.06 & 2.92 \\
  0.75 & 24.81 & 2.85 \\
  1.00 & 19.47 & 3.00 
\end{tabular}
\todo[inline]{Update these values with new periods.  Maybe need to
consider more than one Hamaker constant}
\caption{\label{tbl:aspectRatio} The shear rate, period, and effective
aspect ratio of a vesicle doublet with reduced area $\Delta A = 0.9$,
Hamaker constant $\mathcal{H} = 2$, and separation distance $\delta =
0.4$.  The effective aspect ratios are comparable for all three shear
rates.}
\end{table}
